Airline Customer Satisfaction - Group 17

### 2025-04-02 Draft Business Analytics Assignment Group 17.Rmd
Add files via upload
mohon direview bang

## Data Tranformation <a name="2.4"></a>

```{r}
plot_histogram(data, ncol=2,ggtheme = theme_minimal())
```

We can see that the data is not normally distributed, therefore we need to normalize the data.

@@

plot_histogram(data[, c("Flight.Distance.log", "Age.log", "Departure.Delay.Duration.log", "Arrival.Delay.Duration.log")], ncol=2,ggtheme = theme_minimal())
```

We can see that the data is now normally distributed. Now we need to normalize the data to have the same scale.

@@

@@

Our target is the satisfaction rate of the airlines services, therefore we need to identified how each variable are correlated to the satisfaction

Satisfaction for new features :

@@

```{r}
data %>% group_by(Class, Satisfaction) %>% summarize(num=n()) %>% mutate(percentage=round(num*100/sum(num),2))
```

We notice that economy class passenger is having high customer base but with high percentage of dissatisfied customer rate. Therefore targeting this customer base for services improvement will highly affecting airline overall satisfaction rate.

@@

Personal type of travel passenger is having a third of airline total customer with the number of satisfied customer is only 10% from it. Therefore it is also important customer segment to be improved.

```{r}
data %>% group_by(Customer.Type, Satisfaction) %>%
summarize(num=n()) %>%
mutate(percentage=round(num*100/sum(num),2))
```

76% of first time customer are feels dissatisfied with the airline services, this could be a problem because in the long term it will reducing airlines capability in increasing customer retention.

```{r}
data %>% filter(Customer.Type == 'First-time') %>% group_by(Class ,Satisfaction) %>%
summarize(num=n()) %>%
mutate(percentage=round(num*100/sum(num),2))
```

Most of the first-time customer feel dissatisfied in every class with notably higher in business and economy class.

```{r}
data %>% group_by(Distance_Group) %>%
summarize(num=n()) %>%
mutate(percentage=round(num*100/sum(num),2))
```

Based on the satisfaction rate in each of the distance group, we can notice that as the distance lower the rate of disatisfied customer is increasing. This become a problem because 60% of the airline customer is having journey below 1000.

Next we will try to analyze the service factors related to the customer satisfaction, starting with customer type :

@@

# Plot bar chart with highest value on top and number labels
ggplot(mean_df, aes(x = reorder(Variable, Mean), y = Mean)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = round(Mean, 2)), hjust = -0.2) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Mean of Continuous Variables", x = "Variable", y = "Mean Value")
```

We can identify that from the first time customer, the services that dissapointing them most in Departure and Arrival Time Convenience, Ease of Online Booking, and Wifi Service.

@@

# Plot bar chart with highest value on top and number labels
ggplot(mean_df, aes(x = reorder(Variable, Mean), y = Mean)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = round(Mean, 2)), hjust = -0.2) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Mean of Continuous Variables", x = "Variable", y = "Mean Value")
```

We can identify that from the Personal type of travel customers, the services that dissapointing them most in Departure and Arrival Time Convenience, Ease of Online Booking, and Wifi Service.

@@

# Plot bar chart with highest value on top and number labels
ggplot(mean_df, aes(x = reorder(Variable, Mean), y = Mean)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = round(Mean, 2)), hjust = -0.2) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Mean of Continuous Variables", x = "Variable", y = "Mean Value")
```

We can identify that from the Economy class customers, the services that dissapointing them most in Online Boarding, Ease of Online Booking, and Wifi Service.

### Summary of Initial Plotting

-   43% of total passengers have liked the airline service. The majority of customers who are satisfied are from business class
-   57% of total passengers are not satisfied. The majority of them are from Economy class
-   Only 18% of people have travelled for first time in our airline service. Out of those customers, nearly 76% of customers are not satisfied with our service which could create a chance in future of not returing of customers to our service again. When passengers are coming for first time, most of them have chosen the economy class
-   Out of the customers who opted for business class, nearly 69% of customers are satisfied with the business class. Out of the customers who opted for economy class, nearly 81% of them are not satisfied with it. Customers who opted for economy plus class,nearly 75% of them are not satisfied with it.
-   Whether people are coming for first time or not and also irrespective of their travel type, they are mostly not satisfied with the Economy and Economy Plus classes. There is strong need for the improvement of Economy class since 45% of total passengers are from economy class. Moreover, 60% of passengers travel less than 10k km and most of them prefer economy class.
-   Passengers are not so satisfied with online booking, Online Boarding and the Wifi service mostly.
-   82% of total passengers are returning to the airline service. Out of those 105773 passengers,50% of passengers are from busines class and 42% of passengers are from economy class. Customer who are returning to our service are aged above 30 and make business travel and travel less than 1000 km mostly

>>>>

## Data Tranformation <a name="2.4"></a>
Data transformation plays a key role to make sure all data is in the same scale and distribution. we will start  to take a look on our data distribution.

library(ggplot2)
library(DataExplorer)

# Set default colors for histogram bars
update_geom_defaults("bar", list(fill = "steelblue", color = NA))

# Generate histogram with updated colors
plot_histogram(data, 
               ncol = 2, 
               ggtheme = theme_minimal() + theme(strip.background = element_blank(),
                                                 strip.text = element_text(color = "black")))

```
This part is to undertand the distribution of data before it is used. As we can see in the histogram, the data is not normally distributed. Therefore, we need to normalize the data.
There are several concerning variables in the dataset that are part of main focus in the data transformation process.
<ul>
<li>**arrival.delays** and **departure.delays** have highly-right skewed distribution where most values are concentrated to zero.</li>
<li>**flight.distance** has a right-skewed distribution concentrated between 500 and 1500, with a long tail up to 4000.</li>
<li>**age** has a right-skewed distribution concentrated between 20 and 50, with a long tail up to 80.But this variable is less concerning compared to the previous ones.</li>
</ul>
In the following step we will normalize the data by applying log transformation to the concerning variables.

@@

plot_histogram(data[, c("Flight.Distance.log", "Age.log", "Departure.Delay.Duration.log", "Arrival.Delay.Duration.log")], ncol=2,ggtheme = theme_minimal())
```
Data transformation is performed with visualization task to see the distribution of the data after the transformation. This aims to normalize skewed distribution.
<ul>
  <li>**arrival.delays** and **departure.delays** are transformed into log scale to address the skewed issues.</li>
  <li>**flight.distance** and **age** are also transformed into log scale to address the skewed issues.</li>
</ul>
After log-transormation, normalize the distribution is then performed as follow.

@@

As we can see in the histogram, the data is now normalized with 0 as the center. This is important to ensure that the data is on the same scale and distribution. This helps the model learn the data better and make better predictions.

Here are the reasons why performing normalization is important before using the data:
<ol>
  <li><strong>Comparison Across Variables</strong><br>
      Normalization ensures the data is on the same scale. This helps in making the model comparable and supports easier further analysis.
  </li>
  <li><strong>Improved Model Performance</strong><br>
      Many algorithms perform better when input variables are standardized, as it prevents variables with larger ranges from dominating the analysis.
  </li>
  <li><strong>Normalization Beyond Log Transformation</strong><br>
      While log transformation reduces skewness and normalizes distributions, scaling ensures uniformity across variables, enabling fair use in statistical techniques.
  </li>
</ol>

@@

Our target is the satisfaction rate of the airlines services, therefore we need to identify how each variable are correlated to the satisfaction. This is important to understand how each variable's behavior toward main objective of the analysis.
Following is the behavior of each variables towards satisfaction rate.

```{r}
plot_bar(data,by='Satisfaction',ncol=2)
```
According to the graphic above, we can see that:
<ul>
<li>**Gender**: Both male and female customers have a higher proportion of "Neutral or Dissatisfied" responses compared to "Satisfied" ones. This indicates there is a room of improvement to increase satisfied number</li>
<li>**Customer Type**: Returning customers exhibit a higher proportion of "Satisfied" responses than first-time customers, who seem mostly "Neutral or Dissatisfied". This might indicate that familiarity with the company positively influences satisfaction.</li>
<li> **Type of Travel**: Business travelers show a higher proportion of "Satisfied" responses compared to personal travelers.</li>
<li> **Class**: Business class has the highest proportion of "Satisfied" responses, whereas Economy and Economy Plus lean more towards "Neutral or Dissatisfied." This indicates higher satisfaction among premium customers. </li>
<li> **Age Range**: Older and middle-aged customers show higher satisfaction levels compared to youth and children, who lean towards dissatisfaction.</li>
<li> **Distance Group**:Longer trips seem to lead to higher satisfaction levels, which could indicate that services during these trips meet or exceed customer expectations. </li>
<li> **Delay Status**: the behavior of satisfaction between on-time and delayed status is not significantly different </li>
</ul>

@@

Graphic above is telling the satisfaction distribution based on services category such as comfort, digital, entertainment, and flight services. We can see that the distribution of satisfaction is not significantly different between each category.

```{r}
data %>% group_by(Class, Satisfaction) %>% summarize(num=n()) %>% mutate(percentage=round(num*100/sum(num),2))
```
Based on the data above, **69.44%** of Business passengers showing satisfaction, while only *18.77%* of Economy passengers showing satisfaction and only **24.64%** of Economy Plus passengers are satisfied. This indicates that the airline services are more satisfying for business class passengers compared to economy class passengers. This data highlights the need for airlines to focus on enhancing the customer experience, particularly in Economy and Economy Plus, to achieve better satisfaction rates.

@@

Personal type of travel passenger is having a third of airline total customer with the number of satisfied customer is only **10%** from it. This is much lower compared to the business type customers with **58.37%** satisfaction rate. Personal travel satisfaction levels indicate areas for improvement in service offerings or customer experience. Therefore it is also important customer segment to be improved trying to understand more satisfaction triggers from personal type of travelers.

```{r}
data %>% group_by(Customer.Type, Satisfaction) %>%
summarize(num=n()) %>%
mutate(percentage=round(num*100/sum(num),2))

```

**First-time Customers**: A majority, 76.03%, are neutral or dissatisfied, while only 23.97% report being satisfied.This could be a concern since it is indicating a risk from airline for not retaining first-time customers.<br>
<br>
**Returning Customers**: A majority, 58.37%, are satisfied, while 41.63% are neutral or dissatisfied. This indicates that returning customers are more satisfied with the airline services compared to first-time customers. This data highlights the need for airlines to focus on enhancing the customer experience, particularly for first-time customers, to achieve better satisfaction rates.

```{r}
data %>% filter(Customer.Type == 'First-time') %>% group_by(Class ,Satisfaction) %>%
summarize(num=n()) %>%
mutate(percentage=round(num*100/sum(num),2))

```
The table shows significant differences in first-timer passenger satisfaction across flight classes. Business Class shows the highest satisfaction, with nearly 40% satisfied, though more than half remain neutral or dissatisfied. Economy Class fares worse, with 85% of passengers neutral or dissatisfied, and Economy Plus performs the poorest, with 92% dissatisfaction. Enhancing services for Economy and Economy Plus could greatly improve overall customer satisfaction.<br>
<br>
In conclusion, mostly first-timer passengers feeling dissatisfied on each classes with notably higher in economy and economy plus class. This is a concern for the airline, as it indicates a need to improve services for first-time customers, to retain them and increase satisfaction rates.

```{r}
data %>% group_by(Distance_Group) %>%
summarize(num=n()) %>%
mutate(percentage=round(num*100/sum(num),2))
```
Previously in the barplot satisfaction, we can notice that  as the distance lower the rate of disatisfied customer is increasing. **Below 1000 distance is having the highest dissatisfaction rate**. From the data above, this condition could be a big concern since the majority of the airline customer is having journey below 1000 (more than 60%). This is a concern for the airline, as it indicates a need to improve services for short-distance travelers, to retain them and increase satisfaction rates.
<br>
Next we will try to analyze the service factors related to the customer satisfaction, starting with customer type :

@@

# Plot bar chart with highest value on top and number labels
ggplot(mean_df, aes(x = reorder(Variable, Mean), y = Mean)) +
  geom_bar(stat = "identity", fill = "#CFCFFF") +
  geom_text(aes(label = round(Mean, 2)), hjust = -0.2) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Mean of Continuous Variables", x = "Variable", y = "Mean Value")
```

We can identify that from the first time customer, the services that dissapointing them most in **Departure and Arrival Time Convenience, Ease of Online Booking, and Wifi Service**.While **Comfort Service** having the highest satisfaction value. <br>
This data can guide strategic improvements to enhance the first-time customer experience and increase satisfaction rates by focusing the improvement on those concerning satisfaction factors with the low score.

@@

# Plot bar chart with highest value on top and number labels
ggplot(mean_df, aes(x = reorder(Variable, Mean), y = Mean)) +
  geom_bar(stat = "identity", fill = "#008080") +
  geom_text(aes(label = round(Mean, 2)), hjust = -0.2) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Mean of Continuous Variables", x = "Variable", y = "Mean Value")
```

We can identify that from the Personal type of travel customers, the services that dissapointing them most in **Departure and Arrival Time Convenience, In flight wifi service, Ease of Online Booking**.These areas could be prioritized for enhancement to improve overall satisfaction for business travelers. The top satisfaction factors is **Comfort service**.<br>
As you can see, personal type of travel customer satisfaction data is showing similar behaviour with the first-timer passengers. Thus, improving low-rated areas like WiFi availability and online booking ease may help elevate the overall customer experience for this segment.

@@

# Plot bar chart with highest value on top and number labels
ggplot(mean_df, aes(x = reorder(Variable, Mean), y = Mean)) +
  geom_bar(stat = "identity", fill = "#DC143C") +
  geom_text(aes(label = round(Mean, 2)), hjust = -0.2) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Mean of Continuous Variables", x = "Variable", y = "Mean Value")
```

The graph above is satisfaction data from Economy class customers. The biggest contributors to low satisfaction rate are **ease of online booking, wifi service, and online boarding**. While **Comfort Service** have the highest satisfaction rate among the Economy class customers. This data can guide strategic improvements to enhance the Economy class customer experience and increase satisfaction rates by focusing the improvement on those concerning satisfaction factors with the low score.

### Summary of Initial Plotting
<ol>
<li>43% of total passengers have liked the airline service. The majority of customers who are satisfied are from business class</li>
<li>57% of total passengers are not satisfied. The majority of them are from Economy class </li>
<li>Only 18% of people have travelled for first time in our airline service. Out of those customers, nearly 76% of customers are not satisfied with our service which could create a chance in future of not returing of customers to our service again. When passengers are coming for first time, most of them have chosen the economy class </li>
<li>Out of the customers who opted for business class, nearly 69% of customers are satisfied with the business class. Out of the customers who opted for economy class, nearly 81% of them are not satisfied with it. Customers who opted for economy plus class,nearly 75% of them are not satisfied with it. </li>
<li>Whether people are coming for first time or not and also irrespective of their travel type, they are mostly not satisfied with the Economy and Economy Plus classes. There is strong need for the improvement of Economy class since 45% of total passengers are from economy class. Moreover, 60% of passengers travel less than 10k km and most of them prefer economy class.</li>
<li>Passengers are not so satisfied with online booking, Online Boarding and the Wifi service mostly especially those who take economy class.</li>
<li>82% of total passengers are returning to the airline service. Out of those 105773 passengers,50% of passengers are from busines class and 42% of passengers are from economy class. Customer who are returning to our service are aged above 30 and make business travel and travel less than 1000 km mostly </li>
<li> Departure and arrival time covenience is considered big contributors to the customers dissatisfaction especially for first-timer passengers and personal travel type customers</li>
<li> Passengers mostly take short distance travel (below 1000 km) and it contributes to the high number of dissatisfaction. </li>
</ol>
